---
title: "Seurat Integration"
output: github_document
---

```{r}
install.packages('Seurat')
devtools::install_github('thomasp85/patchwork')
if (!requireNamespace("remotes", quietly = TRUE)) {
  install.packages("remotes")
}

remotes::install_github("satijalab/seurat-wrappers", quiet = TRUE)
remotes::install_github("satijalab/seurat-data", quiet = TRUE)

```

```{r setup, include=FALSE}
library(Seurat)
library(SeuratData)
library(SeuratWrappers)
library(patchwork)
```

##  Read In Data
Seurat v5 assays store data in layers. These layers can store raw, un-normalized counts (layer='counts'), normalized data (layer='data'), or z-scored/variance-stabilized data (layer='scale.data'). 
We can load in the data, remove low-quality cells, and obtain predicted cell annotations (which will be useful for assessing integration later),
 using our Azimuth pipeline.
```{r}
InstallData('pbmcsca')
# load in the pbmc systematic comparative analysis dataset
obj <- LoadData("pbmcsca")
obj <- subset(obj, nFeature_RNA > 1000)
obj <- RunAzimuth(obj, reference = "pbmcref")
obj[["RNA"]] <- split(obj[["RNA"]], f = obj$Method)
obj
```


## Perform Data Without Integration

```{r}
# run standard anlaysis workflow
obj <- NormalizeData(obj)
obj <- FindVariableFeatures(obj)
obj <- ScaleData(obj)
obj <- RunPCA(obj)

obj <- FindNeighbors(obj, dims = 1:30, reduction = "pca")
obj <- FindClusters(obj, resolution = 2, cluster.name = "unintegrated_clusters")

obj <- RunUMAP(obj, dims = 1:30, reduction = "pca", reduction.name = "umap.unintegrated")
# visualize by batch and cell type annotation
# cell type annotations were previously added by Azimuth
DimPlot(obj, reduction = "umap.unintegrated", group.by = c("Method", "predicted.celltype.l2"))
```

Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.
## Perform Integration Through Packages in Seurat
Personally, I have found the most use out of Harmony and FastMNN Integration methods if using Seurat. Always check how the umap integrates data,
we should see a integration of batch effects but retain some sort of biological function (This is a tricky problem and could take additionally analysis to determine the best method for your dataset)

```{r}
obj <- IntegrateLayers(
  object = obj, method = CCAIntegration,
  orig.reduction = "pca", new.reduction = "integrated.cca",
  verbose = FALSE
)

obj <- IntegrateLayers(
  object = obj, method = RPCAIntegration,
  orig.reduction = "pca", new.reduction = "integrated.rpca",
  verbose = FALSE
)

obj <- IntegrateLayers(
  object = obj, method = HarmonyIntegration,
  orig.reduction = "pca", new.reduction = "harmony",
  verbose = FALSE
)

obj <- IntegrateLayers(
  object = obj, method = FastMNNIntegration,
  new.reduction = "integrated.mnn",
  verbose = FALSE
)
```

## Perform Visualization
We can visualize results from both cca and harmony to see how these methods perform on this dataset
```{r}
obj <- FindNeighbors(obj, reduction = "integrated.cca", dims = 1:30)
obj <- FindClusters(obj, resolution = 2, cluster.name = "cca_clusters")

obj <- RunUMAP(obj, reduction = "integrated.cca", dims = 1:30, reduction.name = "umap.cca")
p1 <- DimPlot(
  obj,
  reduction = "umap.cca",
  group.by = c("Method","cca_clusters"),
  combine = FALSE, label.size = 2
)

obj <- FindNeighbors(obj, reduction = "harmony", dims = 1:30)
obj <- FindClusters(obj, resolution = 2, cluster.name = "harmony_clusters")

obj <- RunUMAP(obj, reduction = "harmony", dims = 1:30, reduction.name = "umap.harmony")
p2 <- DimPlot(
  obj,
  reduction = "umap.harmony",
  group.by = c("Method","harmony_clusters"),
  combine = FALSE, label.size = 2
)

wrap_plots(c(p1, p2), ncol = 2, byrow = F)
```