---
title: "BayesPrism Deconvolution"
output: github_document
---
## Install and Load Packages
```{r}
devtools::install_github("Danko-Lab/BayesPrism/BayesPrism")
install.packages('dplyr')
install.packages('Matrix')
install.packages('data.table')

#Install Giotto
# Necessary for installation from R
if(!"devtools" %in% installed.packages()) {
  install.packages("devtools")
}

devtools::install_github("drieslab/Giotto@suite")

library(Giotto)
installGiottoEnvironment()
```
```{r setup, include=FALSE}
library(BayesPrism)
library(Matrix)
library(dplyr)
setwd('~/Desktop/tutorials/Deconvolution/BayesPrism')
```

## Load Reference Dataset and Preprocess

```{r}
#Reference Dataset
ref <- readRDS('../brca_atlas.rds')
sc.mtx <- ref@assays$RNA$counts
metadata <- ref@meta.data
#specify celltypes and cellstates
celltypes <- metadata$celltype_major
cellstates <- metadata$celltype_minor

#Endothelial and PVL have too many states and are too similar in expression
library(data.table)
cellstates[cellstates %like% 'Endothelial'] <- 'Endothelial'
cellstates[cellstates %like% 'PVL'] <- 'PVL'
cellstate_subset <- metadata$celltype_subset

```

### Check Celltype Correlations
we want to see low overlap in different celltypes (i.e., these cells are different enough to call as independent)
Running Deconvolution on similar celltypes usually results in highly innacurate deconvolution
```{r pressure}
sc.mtx <- t(sc.mtx)
plot.cor.phi(input = sc.mtx,input.labels = cellstates,title='cell state correlation',cexRow=0.75,cexCol=0.75,margins=c(8,8))
plot.cor.phi(input=sc.mtx,input.labels = cellstate_subset,title='cell state subset correlations',cexRow=0.75,cexCol=0.75,margins=c(15,15))
plot.cor.phi(input = sc.mtx,input.labels = celltypes,title='cell type correlation',cexRow=1.0,cexCol=1.0)
```

### Filter Outlier Genes and Subset to Protein Coding
When running deconvolution, it is imperative to utilize good genes that differentiate celltypes, are unique to each cell, have adequate expression.
Without good genes, we will not be able to get a good results
Typically this is the most important step, most tools perform well if good data is used
```{r}
sc.stat <- plot.scRNA.outlier(input = sc.mtx,cell.type.labels = cellstates,species = 'hs',return.raw = TRUE)
sc.mtx.filtered <- cleanup.genes(input = sc.mtx,input.type = 'count.matrix',species = 'hs',
                                 gene.group = c('Rb','Mrp','MALAT1','chrM','chrX','chrY'),
                                 exp.cells = 5)

sc.mtx.filtered.pc <- select.gene.type(sc.mtx.filtered,gene.type = 'protein_coding')
```

### Subset Genes to Differential Genes
Not essential, but could help when celltypes are similar
This analysis uses differential gene expression, sets cutoffs for the genes found, and subsets object to these genes
Adjust the thresholding if too few genes are found (Personally, I like using at least 100 genes per celltype)
```{r}
diff.exp.stat <- get.exp.stat(sc.dat = sc.mtx[,colSums(as.matrix(sc.mtx)>0)>3],
                              cell.type.labels = cellstates,
                              cell.state.labels = cellstate_subset,
                              pseudo.count = 0.1,
                              cell.count.cutoff = 50,
                              n.cores=1)
sc.mtx.filtered.pc.sig <- select.marker(sc.dat = sc.mtx.filtered.pc,
                                        stat=diff.exp.stat,
                                        pval.max = 0.01,
                                        lfc.min = 0.1)
```

# Run BayesPrism
```{r}
library(Giotto)
readRDS('../GV25999_mouse.rds') -> vehicle.mouse
readRDS('../GV26001_mouse.rds') -> ac21.mouse
readRDS('../GV26002_mouse.rds') -> ac50.mouse

vehicle.expr <- t(vehicle.mouse@expression$cell$rna$raw@exprMat)
ac21.expr <- t(ac21.mouse@expression$cell$rna$raw@exprMat)
ac50.expr <- t(ac50.mouse@expression$cell$rna$raw@exprMat)
```

### Construct BayesPrism Object on Reference Dataset
```{r}
myPrism <- new.prism(
  reference=as.matrix(sc.mtx.filtered.pc.mouse),
  mixture = as.matrix(vehicle.expr),
  input.type = 'count.matrix',
  cell.type.labels = celltypes,
  cell.state.labels = cellstate_subset,
  key = NULL,
  outlier.cut = 0.1,
  outlier.fraction = 0.1
)
```

### Run BayesPrism
This step can take an enourmous amount of time and I suggest running on the cluster and inclreasing  the n.cores to >20 
```{r}
veh.res <- run.prism(prism = myPrism, n.cores=4)

#Extract the Posterior Mean of Cell Type Fraction Theta
veh.theta <- get.fraction (bp=veh.res,
                                    which.theta="final",
                                    state.or.type="type")

#Extract Coefficient of Variation (CV) of Cell Type Fraction
theta.cv <- veh.res@posterior.theta_f@theta.cv

```

**Downstream analysis**
Potential downstream analysis can be performed using theta and Z are:

- Clustering bulk samples by theta or Z (Z can be normalized by vst(round(t(Z.tumor))), using the vst function from the DESeq2 package.)

- Computing z-scores of signature genes for Z of the cell type of interest.

- Survival analysis and correlation with other clinical covariates using theta and Z.

- Correlating Z (after normalization using vst or from bp.res@reference.update@psi_mal) with theta to understand how gene expression of each gene (in malignant cells) correlates with the cell type fraction of non-malignant cells in tumor micro-environment, followed by gene set enrichment analysis (as done in BayesPrism paper).

- Embedding learning of malignant gene expression (see “Tutorial: embedding learning of malignant cell expression using BayesPrism”)
