---
title: "Colocalization"
output: github_document
---
# Analyze Spatial Correlations Between Two Different Groups
In this tutorial we will utilize cell type deconvolution and PAGE Enrichment dataframes to determine the amount of correlation between these two results.
This analysis gives us potential information on how tumor/stroma+immune interact and influence one another

```{r setup, include=FALSE}
library(Giotto)
library(data.table)
library(dplyr)
library(corrplot)
```


## Load in Data and Visualize
This is preprocessed visium dataset through Giotto, which has the following preprocessing performed on it:
1. Filtering
2. Normalization
3. Dimensionality reduction
4. Clustering
5. Cell Type Deconvolution 
6. PAGE Enrichment Test

```{r cars}
setwd('~/Desktop/tutorials/Visium/Colocalization')
readRDS('../data/GV26001_mouse.rds') -> mouse 
readRDS('../data/GV26001_human.rds') -> human
```

Visualize cell type information
```{r}
# Set as dataframe object, set rownames, and remove column cell_id
cell_type <- mouse@spatial_enrichment$cell$rna$RCTD@enrichDT %>% as.data.frame()
rownames(cell_type) <- cell_type$cell_ID
cell_type <- cell_type %>% select(!cell_ID)
head(cell_type)
```

Visualize tumor enrichment results (PAGE)
```{r}
tumor_enrich <- human@spatial_enrichment$cell$rna$PAGE@enrichDT %>% as.data.frame()
rownames(tumor_enrich) <- tumor_enrich$cell_ID
tumor_enrich <- tumor_enrich %>% select(!cell_ID)
tumor_enrich <- tumor_enrich[,colnames(tumor_enrich) %like% 'tumor_enrich'] #just look at subset of results
head(tumor_enrich)
```

## Visualize Spatial Plots
```{r}
library(RColorBrewer)
SpatialColors <- colorRampPalette(colors = rev(x = brewer.pal(n = 11, name = "Spectral")))
FeaturePalettes <- list(
  'Spatial' = SpatialColors(n = 100),
  'Seurat' = c('lightgrey', 'blue')
)
```

```{r}
#Choose Cell Types to Visualize
visualize <- colnames(cell_type)[1:10]
spatCellPlot2D(mouse,
    cell_annotation_values = visualize,
    spat_enr_names = 'RCTD',
    cell_color_gradient = FeaturePalettes$Spatial,
    cow_n_col = 2)
```

```{r}
#Choose Cell Types to Visualize
visualize <- colnames(tumor_enrich)[1:10]
spatCellPlot2D(human,
    cell_annotation_values = visualize,
    spat_enr_names = 'PAGE',
    cell_color_gradient = FeaturePalettes$Spatial,
    cow_n_col = 2)
```

## Correlation Analysis
```{r}
#set cell id to be ordered in the same way
tumor_enrich <- tumor_enrich[rownames(cell_type),]

#make dataframes numeric
cell_type <- apply(cell_type,2,as.numeric)
tumor_enrich <- apply(tumor_enrich,2,as.numeric)

#Run Correlation Analysis
correlation <- cor(tumor_enrich,cell_type)
head(correlation)
```
