---
title: "Niches Cell-Cell Interaction"
output: github_document
---

```{r}
install.packages('Seurat')
install.packages('ggplot2')
install.packages('viridis')
install.packages('dplyr')
devtools::install_local('NICHES-1.1.0.zip')
install.packages("remotes")
remotes::install_github("satijalab/seurat-wrappers")
```

# Perform Cell Cell Interaction Through Matrix Methods
In this example, we utilize a method that fundamentally multiplies a ligand with receptor expression in a spatial neighborhood. Steps Include:
1. Loading Data
2. Creating Custom Database to Investigate Human/Mouse Interactions (different organisms)
3. Impute Gene Expression to remove high zeros in matrix
4. Create a Spatial Neighborhood
5. Perform Neighborhood To Cell Matrix Multiplication
6. Create New Seurat Object and Visualize results
7. Perform Clustering and Differential Gene Expression

```{r setup, include=FALSE}
library(Seurat)
library(ggplot2)
library(dplyr)
library(SeuratWrappers)
library(NICHES)
library(viridis)
setwd('~/Desktop/tutorials/Visium/CCI')
```



## Load Data 
```{r}
setwd('~/Desktop/tutorials/Visium/CCI')
#READ IN MTX FILE
human <- read.table('../data/GV26001_human_counts.tsv.gz',sep='\t',header=TRUE,row.names=1)
mouse <- read.table('../data/GV26001_mouse_counts.tsv.gz',sep='\t',header=TRUE,row.names = 1)
#Subset on in tissue barcodes
tissue <- read.table('../data/spatial/tissue_positions_list.csv',sep=',',row.names=1)
in_tissue <- tissue[tissue$V2 == 1,]
rownames(in_tissue) -> keeps

#combine datasets
human <- human[,colnames(mouse)]
counts <- rbind(human,mouse)
counts <- counts[,keeps] #subset on in tissue
# MAKE SEURAT OBJECT
x <- CreateSeuratObject(counts,assay = 'Spatial',project = 'CCI')
```
```{r}
# READ IN IMAGE FILE
img = Read10X_Image(image.dir = '../data/spatial',image.name = 'tissue_lowres_image.png',assay='Spatial',slice='slice1',filter.matrix = TRUE)
# ADD IMAGE TO SEURAT OBJECT
x@images$image <- img
```


## Create Custom Database
This is an adjustment of normal ligand (human) to receptor (human) or vise versa, we are mixing source and target organisms for our unique dataset
- **DON"T** do this IF the dataset isn't a xenograft model
```{r}
custom_database <- function(){
  ground.truth.human <- LoadFantom5('human')
  ground.truth.mouse <- LoadFantom5('mouse')
#Make Custom Databases for Dataset (mix mouse and human)
  human.to.mouse <- list('source.subunits' = ground.truth.human$source.subunits,'target.subunits' = ground.truth.mouse$target.subunits)
  mouse.to.human <- list('source.subunits' = ground.truth.mouse$source.subunits,'target.subunits' = ground.truth.human$target.subunits)
  human.human <- list('source.subunits' = ground.truth.human$source.subunits,'target.subunits' = ground.truth.human$target.subunits)
  mouse.mouse <- list('source.subunits' = ground.truth.mouse$source.subunits,'target.subunits' = ground.truth.mouse$target.subunits)
  ground.truth <- list('source.subunits' = rbind(human.to.mouse$source.subunits,mouse.to.human$source.subunits),
                       'target.subunits' = rbind(human.to.mouse$target.subunits,mouse.to.human$target.subunits))
  return(ground.truth)
}

db <- custom_database()
db <- as.data.frame(db)
colnames(db) <- c('source','target')
db <- db[!duplicated(db),]
```


## Perform Basic Analysis
```{r cars}
x <- NormalizeData(x, verbose = FALSE)
x <- FindVariableFeatures(x)
x <- ScaleData(x)
x <- RunPCA(x, verbose = FALSE)
x <- FindNeighbors(x, dims = 1:30)
x <- FindClusters(x, verbose = FALSE)
x <- RunUMAP(x, dims = 1:30)

coordinates <- GetTissueCoordinates(x)
x@meta.data$x <- coordinates$x
x@meta.data$y <- coordinates$y


#RUN IMPUTATION ON DATASET
x <- SeuratWrappers::RunALRA(x)
```

## Compute Ligand/Receptor Matrix using package NICHES
```{r pressure, echo=FALSE}
#compute edgelist (which cells are connected)
edgelist <- compute_edgelist(x,position.x = 'x',position.y = 'y',k=4,rad.set = NULL)
#Filter database by genes present in matrix
ground_truth <- FilterGroundTruth(db,input_rownames = rownames(x@assays$alra@data))
#RUN Matrix calculations and generate CCI 
NICHES_output <- RunNICHES(object = x,
                           LR.database = 'custom',
                           custom_LR_database =db,
                           output_format = 'seurat',
                           assay = "alra",
                           edgelist = edgelist,
                           position.x = 'x',
                           position.y = 'y',
                           meta.data.df = x@meta.data,
                           CellToCell = F,CellToSystem=F,SystemToCell=F,CellToCellSpatial=F,CellToNeighborhood=F,NeighborhoodToCell=T)
```

We next isolate this output and embed using UMAP to visualize the microenvironemnt of each cell.
```{r}
niche <- NICHES_output[['NeighborhoodToCell']]
Idents(niche) <- niche[['ReceivingType']]

# Scale and visualize
niche <- ScaleData(niche)
niche <- FindVariableFeatures(niche,selection.method = "disp")
niche <- RunPCA(niche)
ElbowPlot(niche,ndims = 50)
```

```{r}
niche <- RunUMAP(niche,dims = 1:10)
DimPlot(niche,reduction = 'umap',pt.size = 0.5,shuffle = T, label = T) +ggtitle('Cellular Microenvironment')+NoLegend()
```

## Differential Analysis
```{r}
# Find markers
mark <- FindAllMarkers(niche,min.pct = 0.25,only.pos = T,test.use = "roc")
GOI_niche <- mark %>% group_by(cluster) %>% top_n(5,myAUC)
DoHeatmap(niche,features = unique(GOI_niche$gene))+ 
  scale_fill_gradientn(colors = c("grey","white", "blue"))
```

## Make Spatial Plots To Visualize LR Activity
Adjust genes used for ones you are interested in 
```{r}
# Check that these make sense and print spatial Plots
DefaultAssay(x) <- 'alra'
p1 <- SpatialFeaturePlot(x, crop = TRUE, features = "PDGFA",slot = "data",min.cutoff =  'q1',
                         max.cutoff = 'q99')+ggtitle("Ligand")+theme(legend.position = "right")
p2 <- SpatialFeaturePlot(x, crop = TRUE, features = "Pdgfrb",slot = "data",min.cutoff =  'q1',
                         max.cutoff = 'q99')+ggtitle("Receptor")+theme(legend.position = "right")

ggpubr::ggarrange(p1,p2)
```

## More Interestingly, Plot The Combined Expression Spatially on New Object
```{r}
# Add Niches output as an assay
niches.data <- GetAssayData(object =  niche[['NeighborhoodToCell']], slot = 'data')
colnames(niches.data) <- niche[['ReceivingCell']]$ReceivingCell
x[["NeighborhoodToCell"]] <- CreateAssayObject(data = niches.data )
DefaultAssay(x) <- "NeighborhoodToCell"
x <- ScaleData(x)
SpatialFeaturePlot(x,crop=TRUE,features = "PDGFA—Pdgfrb")
```